Coming to Germany: Language, Job search and Labor market outcomes

### 2018-05-21 Paper 2.tex
Update Paper 2.tex

Do people with some host-country language skills at arrival integrate better in the long-run than the ones who do not? Since the 2000s, six Western European countries \footnote{Austria, Denmark, France, Germany, the Netherlands, the UK} have laid down the law that requires pre-departure integration conditions for all non-EU family migrants (Bonjour, 2014). Specifically, since August 2007, foreign adults migrating as spouses of German nationals have to demonstrate at least basic level of German language skills, while foreign children between 16 and 18 years old have to demonstrate advanced level of German language skills, sufficient to start learning in high schools or vocational schools (Block, 2012). If this pre-departure integration requirement makes sense for immigrants coming to Germany for educational or working reasons, does that make equally sense for family reunion purpose?

@@

This study attempts to answer these questions by investigating the role of the pre-migration language skills on the contemporary labor market outcomes of immigrants in Germany, in particular, the earnings and the quality of jobs. However, unlike many existing empirical studies which mainly reports the 'total effect' of language fluency on a range of labor market outcomes, I attempt to open the black-box to disentangle \textbf{how} and \textbf{why} language proficiency exerts its influence on labor market success of immigrants.

Chiswick and Miller (2005) establish that language skills, especially at the beginning of migration, facilitate early job search and the establishment of social networks. At arrival to the host country, immigrants face with various choices of searching for jobs, and they weight different alternatives by their relative costs (Beggs and Chapman, 1990; Frijter et al. 2003; Battu et al., 2011). Immigrants possessing some level of language fluency have lower costs to embrace the mainstream job search methods such as public and private employment services or job advertisements on public media. The ones with poor language skills find it relatively less costly to search jobs via their own informal networks. Many studies in a range of countries find that the job search methods actually matter for the labor market performance. In particular, Wahba and Zenou (2005); Bentolia et al. (2010); Pellizzari (2010) show that using personal network of friends and family is a popular and effective search channel to obtain job offers. But the effectivessness of this search method in terms of later earnings and job quality is questionable as it depends on the quality of an individual's social network composition (Battu et al., 2011; Cappellari and Tatsiramos, 2015).

>>>>

Do people with some host-country language skills at arrival integrate better in the long-run than the ones who do not? Perhaps yes, otherwise, many Western European countries \footnote{Austria, Denmark, France, Germany, the Netherlands, the UK} would not have issued a so-called pre-departure integration condition "to promote integration" for all non-EU family migrants. Specifically, since August 2007, foreign adults migrating as spouses of German nationals have to demonstrate at least basic level of German language skills, while foreign children between 16 and 18 years old have to demonstrate advanced skills of this language (Block, 2012). If this pre-departure integration requirement is essential for immigrants coming to Germany for educational or working reasons, is that also necessary for family reunion purpose?

@@

Leaving this question aside for a moment and consider another well-established strand of literature - job search methods of immigrants. Many studies in a range of countries find that the job search methods actually matter for the labor market performance. In particular, Wahba and Zenou (2005); Bentolia et al. (2010); Pellizzari (2010) show that using personal network of friends and family is a popular and effective search channel to obtain job offers. But the effectivessness of this search method in terms of later earnings and job quality is questionable as it depends on the quality of an individual's social network composition (Battu et al., 2011; Cappellari and Tatsiramos, 2015).

This study combines these two strands of literature and hypothesizes that the pre-migration language skills has an impact on the choice of early job search, which in turn affects the long-run labor market outcomes of immigrants in Germany. Unlike many existing empirical studies which mainly reports the 'total effect' of language fluency on a range of labor market outcomes, I attempt to open the black-box to disentangle \textbf{how} and \textbf{why} language proficiency exerts its influence on labor market success of immigrants. Using mediation analysis

% Chiswick and Miller (2005) establish that language skills, especially at the beginning of migration, facilitate early job search and the establishment of social networks. At arrival to the host country, immigrants face with various choices of searching for jobs, and they weight different alternatives by their relative costs (Beggs and Chapman, 1990; Frijter et al. 2003; Battu et al., 2011). Immigrants possessing some level of language fluency have lower costs to embrace the mainstream job search methods such as public and private employment services or job advertisements on public media. The ones with poor language skills find it relatively less costly to search jobs via their own informal networks.
